### 2024-02-07 products/manuscript/Manuscript.qmd
Project Part 1

---
title: ""
subtitle: ""
author: Mutsa Nyamuranga
date: "02/01/2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

# Summary/Abstract

{{< pagebreak >}}

# Introduction

## General Background Information

Atherosclerosis, a chronic inflammatory disease of the arterial wall, remains a leading cause of cardiovascular morbidity and mortality worldwide. Despite significant advancements in understanding its pathophysiology, effective therapeutic interventions are still in demand. Recent research has shed light on the potential role of dietary β-carotene and its conversion to vitamin A in modulating atherosclerosis progression and resolution.

Several studies have demonstrated the beneficial effects of β-carotene in delaying atherosclerosis progression in experimental models, along with its association with reduced cardiovascular risk in humans. Moreover, genetic variants in the BCO1 gene, responsible for converting β-carotene to vitamin A, have been linked to alterations in circulating carotenoid levels and plasma lipid profiles in individuals. However, the precise mechanisms underlying these effects, particularly regarding the interplay between β-carotene, BCO1 activity, and atherosclerosis progression or resolution, remain unclear.

## Description of data and data source
The data I will be using is based on:

Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.

The data consists of measurements of circulating β-carotene and retinoic acid levels, along with relevant biomarkers and clinical characteristics, collected from a cohort of individuals with varying degrees of atherosclerosis progression and resolution. Genetic information on BCO1 activity, including genetic variants associated with BCO1 activity, may also be included in the dataset. The data was collected through blood sample analysis and genetic testing.

```{r}
# Load necessary packages
library(renv)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths

# Load the dataset (replace "file_path" with the actual file path)
# Load necessary package
library(readxl)

# Read Excel file
rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
```

## Questions/Hypotheses to be addressed

  The question I want to answer using the data is whether there is a correlation between circulating β-carotene levels, BCO1 genetic variants, and the progression or resolution of atherosclerosis. Specifically, I am interested in examining how these factors relate to plasma lipid profiles (e.g., cholesterol, triglycerides) and immune cell populations, particularly Treg levels, in individuals with atherosclerosis. 
  The main outcome of interest will be the degree of atherosclerosis progression or resolution, potentially measured by imaging techniques or clinical assessments of plaque burden. Additionally, plasma lipid profiles and immune cell populations, particularly Treg levels, would be of interest as secondary outcomes. The specific predictors of interest would include circulating β-carotene levels, genetic variants associated with BCO1 activity, plasma lipid profiles, and immune cell populations.
  The data analysis will likely involve statistical methods such as correlation analysis to assess the relationships between predictors and outcomes. Regression modeling could be used to examine the predictive value of circulating β-carotene levels and BCO1 genetic variants on atherosclerosis progression or resolution, controlling for potential confounding variables. Further exploration may involve visualization techniques to identify patterns and trends in the data.

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2024-03-10 R/analysis-code/exploratory-analysis.qmd
Part 2

---
title: "Data Exploration"
author: "Mutsa Nyamuranga"
date: "2024-02-23"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files. And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyr)
library(forcats)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
figdata <- here::here("data","processed_data","processeddata.rds")
#load data
explorfigdata <- readRDS(figdata)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_explorfigdata = skimr::skim(explorfigdata)
print(summary_explorfigdata)
# save to file
summarytable_elife = here("results", "summarytable.rds")
saveRDS(summary_explorfigdata, file = summarytable_elife)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Here, I have created code that develops a bar graph to visualize the distribution of body weight in each condition. This will show the essential volume of body weight throughout the 16 weeks. This will not not be reliable for inferences but essential for drawing a baseline.

## Total Weights

```{r}
# Convert columns to numeric if needed
explorfigdata <- mutate_all(explorfigdata, as.numeric)

# Remove the "Week" variable
explorfigdata1 <- select(explorfigdata, -Week)

# Reshape the data and calculate sum for each condition
explorfigdata_long <- explorfigdata1 %>%
  gather(condition, value) %>%
  filter(!is.na(value)) %>%  # Remove NA values if any
  group_by(condition) %>%
  summarise(sum_value = sum(value))

# Create Bar Graph for Visualization
bar1 <- ggplot(explorfigdata_long, aes(x = fct_reorder(condition, desc(sum_value)), y = sum_value, fill = condition)) +
  geom_col() +
  scale_fill_manual(values = c("red", "blue", "blue", "red", "red", "blue")) +  # Fill colors
  scale_color_manual(values = c("red", "blue", "blue", "red", "red", "blue")) +  # Border colors
  labs(title = "Total Body Weight for Conditions",
       x = "Condition", y = "Sum of Weights")

plot(bar1)
# Saving the plot as an image file
figure_file = here("results","condition_distro.png")
ggsave(filename = figure_file, plot=bar1) 
```

I color coded for male and female stratified conditions. Here, we see that the beta-carotene condition for males has the highest overall total body weight, while the baseline female condition has the lowest weight.

## Weight over time

```{r}
# Reshape data to long format
library(tidyr)
body_weights_long <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")

summary(body_weights_long)
str(body_weights_long)

# Plot the data
library(ggplot2)
ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
  geom_line() +
  labs(title = "Comparison of Body Weight Changes Across Categories",
       x = "Week", y = "Body Weight") +
  theme_minimal()
```

With some final exploration. We look at the change in body weight over the 16 weeks. We see that males start and finish with higher body weights as expected. Of note, we see that the control and Beta-Carotene condition for female mice ends with the lowest body weights.

### 2024-03-10 R/processing-code/processingfile-v1.qmd
Part 2

---
title: "Data Processing"
author: "Mutsa Nyamuranga"
date: "2024-02-23"
output: html_document
---

# Setup
In terms of packages, I need to load these set in order to view, explore and clean the raw data from the eLife paper.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

With difficulty using the here package for creating a path to the excel document, I have chosen to use the readxl package to directly import the file into the environment.

```{r}
library(readxl)
rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
```

# Check data

```{r}
library(readxl)

range_to_read <- "A4:M24"

# Read the Excel file into a data frame
figure1a <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx", sheet = "Figure 1-suppl figure 1", range = range_to_read)

summary(figure1a)
```

#Clean Data

Here I am taking the Figure 1a data table from the paper and clean the data to fit my purposes for analysis. I remove the standard deviation data and fix female and male onto each condition type.

```{r}
# Remove specified columns and rows
figure1a<- figure1a[, -c(3, 5, 7, 9, 11, 13)]
figure1afixed <- figure1a[-1, ]

# Fixing Female and Male to condition
figdata <- figure1afixed %>%
  rename(BaselineF = Baseline...2,
         ControlF = Control...4,
         BetaCaroteneF = `β-carotene...6`,
         BaselineM = Baseline...8,
         ControlM = Control...10,
         BetaCaroM = `β-carotene...12`)

# View the updated data frame
print(figdata)
str(figdata)
summary(figdata)
glimpse(figdata)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
dir.create("/Users/mutsa_n/Desktop/MADA-course/Nyamuranga-MADA-project/data/processed_data/", recursive = TRUE)
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(figdata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-03-10 products/manuscript/Manuscript.qmd
Part 2

@@
 
-# Load the dataset (replace "file_path" with the actual file path)
 # Load necessary package
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-
 ## Schematic of workflow
@@
 ## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+The data was acquired online from Pinos et.
+al.
+in the additional files section.
+The data was formated in tables within an excel file which had to be processed and cleaned to identify target data.
+The excel file was loaded in through readxl and downloaded into a processing quarto file.
 
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+
+The provided R code is structured to process and clean raw data extracted from an Excel file.
+Initially, the data is loaded into R using the read_excel function, and a specific range of cells is extracted from the Excel sheet "Figure 1-suppl figure 1".
+The subsequent cleaning steps involve removing columns containing standard deviation data and ensuring clarity and consistency by labeling conditions with "Female" and "Male" appropriately.
+Following this, basic exploration functions such as print, str, summary, and glimpse are utilized to examine the cleaned data frame's structure and content.
+Finally, the cleaned data is saved as an RDS file for future analysis.
 

### 2024-03-15 R/processing-code/processingfile-v1.qmd
Fix on loading data

@@
 ```{r}
-library(readxl)
-rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
+# path to data
+# note the use of the here() package and not absolute paths
+data_location <- here::here("data/raw-data/elife-87430-data1-v1.xlsx")
+rawdata <- readxl::read_excel(data_location)
 ```
@@
 ```{r}
-library(readxl)
-
 range_to_read <- "A4:M24"
 
-# Read the Excel file into a data frame
-figure1a <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx", sheet = "Figure 1-suppl figure 1", range = range_to_read)
+figure1a <- readxl::read_excel(data_location, sheet ="Figure 1-suppl figure 1" , range = range_to_read)
 
-summary(figure1a)
+print(figure1a)
 ```

### 2024-03-16 R/analysis-code/exploratory-analysis.qmd
fix 2

@@
 ---
-
-This Quarto file loads the cleaned data and does some exploring.
-
-I'm only showing it the way where the code is included in the file.
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
-
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
 

### 2024-03-18 products/manuscript/Manuscript.qmd
Part 3

@@
 
-## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("products","assets","antigen-recognition.png"))
-```
-
 ## Data aquisition
@@
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
+
+```{r load_image, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
+library(png)
+condimage <- readPNG(here("results", "figures", "condition_distro.png"))
+par(mar = c(1, 1, 1, 1))
+plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
+rasterImage(condimage, 0.1, 0.1, 0.9, 0.9)  
+```
+
+```{r load_image, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
+library(png)
+catimage <- readPNG(here("results", "figures", "category_distro.png"))
+par(mar = c(1, 1, 1, 1))
+plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
+rasterImage(catimage, 0.1, 0.1, 0.9, 0.9)  
+```
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
 @tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
 
@@
 #| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
+summarytable = readRDS(here("results","tables","summarytable.rds"))
+knitr::kable(summarytable)
 ```
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+## Full analysis
 
 ```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| label: linreg
+#| tbl-cap: "Linear model fit table."
 #| echo: FALSE
-knitr::include_graphics(here("results","height-weight-stratified.png"))
+resulttable1 = readRDS(here("results","output","resulttable1.rds"))
+knitr::kable(resulttable1)
+```
+Linear Regression: The results reveal the estimated effects of different treatment categories on the response variable, body weight, along with their statistical significance.
+The intercept, estimated at 25.0, represents the baseline level of the response variable when all predictors are zero.
+Relative to a reference category, presumably "BaselineM", coefficients associated with each treatment category provide insights into their effects on the response variable.
+For instance, "BaselineM" and "BetaCaroM" are associated with increases in the response variable by approximately 8.63 and 9.85 units, respectively, compared to the reference category.
+Conversely, "BetaCaroteneF", "ControlF", and "ControlM" exhibit smaller or even negative effects, but the statistical significance varies among these categories, with "BetaCaroteneF" showing a non-significant effect and others demonstrating statistically significant effects with low p-values.
+
+```{r}
+#| label: timeseriestable
+#| tbl-cap: "Time Series model fit table."
+#| echo: FALSE
+resulttable2 = readRDS(here("results","output","resulttable2.rds"))
+knitr::kable(resulttable2)
 ```
 
-## Full analysis
+Time Series Model: The linear regression model was constructed to investigate the relationship between body weight and time (Week) while considering treatment variables.
+The model revealed several key findings.
+Firstly, the estimated intercept of 14.9634 suggests that the average body weight at the beginning of the observation period, when all other predictors are zero, is approximately 14.9634 units.
+Additionally, the coefficient for Week, estimated at 1.1047, indicates that body weight tends to increase by approximately 1.1047 units for each additional week, assuming a linear trend.
+However, the quadratic and cubic terms of Week were not statistically significant, suggesting that the relationship between body weight and time may not exhibit quadratic or cubic trends.
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
+Furthermore, the model uncovered notable effects of treatment variables on body weight.
+Specifically, compared to the reference level "BaselineM", various treatment categories exhibited significant impacts on body weight.
+For instance, "BetaCaroM" and "ControlM" categories were associated with increased body weight by 8.6331 and 8.1861 units, respectively, compared to "BaselineM".
+Conversely, "BetaCaroteneF" and "ControlF" categories were associated with decreases in body weight by 1.5694 and 1.0445 units, respectively, compared to "BaselineM".
+These findings collectively suggest that both time and treatment variables play crucial roles in influencing body weight variations over the observed period.
 

### 2024-03-18 R/analysis-code/statistical-analysis.R

@@ -5,16 +5,23 @@
 #and saves the results to the results folder
 
 #load needed packages. make sure they are installed.
+library(here) #for data loading/saving
+library(dplyr)
+library(skimr)
+library(ggplot2)
+library(tidyr)
+library(forcats)
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
 
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+#Load the data.
 
-#load data. 
-mydata <- readRDS(data_location)
+#Path to data. Note the use of the here() package and not absolute paths
+figdata <- here::here("data","processed_data","processeddata.rds")
+#load data
+explorfigdata <- readRDS(figdata)
+body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
 
 
 ######################################
@@ -25,32 +32,40 @@ mydata <- readRDS(data_location)
 #### First model fit
 # fit linear model using height as outcome, weight as predictor
 
-lmfit1 <- lm(Height ~ Weight, mydata)  
+linCat <- lm(BodyWeight ~ Category, body_weights)  
 
 # place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
+linCattable <- broom::tidy(linCat)
 
 #look at fit results
-print(lmtable1)
+print(linCattable)
 
 # save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+table_linCat = here("results", "resulttable1.rds")
+saveRDS(linCattable, file = table_linCat)
 
 ############################
 #### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
 
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
+# Assuming your time series data is stored in your_data with Week as the time variable and BodyWeight as the data
+
+# Load required library
+library(strucchange)
+
+# Fit a linear regression model with treatment variables
+timeseries <- lm(BodyWeight ~ Week + I(Week^2) + I(Week^3) + Category, data = body_weights)
+
+# Summary of the model
+summary(timeseries)
 
 # place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
[…]

### 2024-04-09 R/analysis-code/modeling.qmd
Project 4

---
title: "Model Fitting"
author: "Mutsa Nyamuranga"
date: "2024-03-29"
output: html_document
---

# Introduction

# Set Up

## Load Necessary Packages
First we will load the data and additional packages necessary for the
assigniment.

```{r}
library(readr) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(ggplot2) # for plots
library(gtsummary)# for summary tables
library(patchwork) #for combine plots
library(tidymodels)
library(recipes)
library(parsnip)
```

## Load Data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
figdata <- here::here("data","processed_data","processeddata.rds")
#load data
explorfigdata <- readRDS(figdata)
body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
```

# Modeling

Here, I conduct a model fitting to assess the relationship of the
outcome of interest with the other variables. Tidymodels provides the
functions necessary for modeling and preprocessing data. The recipe
function specifies the data preprocessing steps. In this case, the
outcome variable (Y) is defined as the response variable, and all other
variables are considered predictors. step_dummy converts categorical
predictors into dummy variables, step_center centers numeric predictors
around their mean, and step_scale scales numeric predictors to have unit
variance.

## First model fit
```{r}
# fit linear model using height as outcome, weight as predictor

linCat <- lm(BodyWeight ~ Category, body_weights)

# place results from fit into a data frame with the tidy function
linCattable <- broom::tidy(linCat)

#look at fit results
print(linCattable)

# save fit results table  
table_linCat = here("results", "resulttable1.rds")
saveRDS(linCattable, file = table_linCat)
```

## Second model fit
```{r}
# Load required library
library(strucchange)

# Fit a linear regression model with treatment variables
timeseries <- lm(BodyWeight ~ Week + I(Week^2) + I(Week^3) + Category, data = body_weights)

# Summary of the model
summary(timeseries)

# place results from fit into a data frame with the tidy function
timeseriestbl <- broom::tidy(timeseries)

#look at fit results
print(timeseriestbl)

# save fit results table  
seriestable_file2 = here("results", "resulttable2.rds")
saveRDS(timeseriestbl, file = seriestable_file2)
```
## Third Model Fit
```{r}
# Define the recipe
body_recipe <- recipe(BodyWeight ~ ., data = body_weights) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
```

```{r}
# Split the data into training and testing sets
set.seed(123) # for reproducibility
body_split <- initial_split(body_weights, prop = 0.8, strata = BodyWeight)
body_train <- training(body_split)
body_test <- testing(body_split)

# Define the linear regression model specification
linear_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Create a workflow that incorporates the recipe and the model
body_workflow <- workflow() %>%
  add_model(linear_spec) %>%
  add_recipe(body_recipe)
```

The initial_split function divides the dataset into training and testing
sets. The prop argument specifies the proportion of the data to allocate
to the training set (80% in this case), and the strata argument ensures
that the split is stratified based on the outcome variable (Y), which
helps maintain the balance of categories in both sets.

The linear_reg function specifies the linear regression model.
set_engine("lm") selects the linear model engine (ordinary least squares
regression), and set_mode("regression") sets the mode of the model to
regression, indicating that it predicts a continuous outcome (Y).

The workflow function creates a modeling workflow. This workflow
combines the preprocessing steps defined in the recipe with the
specified model. The add_recipe function adds the recipe to the
workflow, and the add_model function adds the linear regression model.

```{r}
body_weights$BodyWeight <- as.numeric(as.character(body_weights$BodyWeight))

class(body_weights$BodyWeight)
```

```{r}
# Fit the workflow to the training data
body_fit <- fit(body_workflow, data = body_train)

# Compute predictions on the testing data
body_predictions <- predict(body_fit, new_data = body_test) %>%
  bind_cols(body_test)
```

The fit function fits the workflow to the training data. This step
applies the preprocessing steps defined in the recipe to the training
data and then fits the linear regression model to the preprocessed data.

The predict function generates predictions from the fitted model using
the testing data. This step applies the preprocessing steps (from the
recipe) to the testing data and then predicts the outcome variable (Y)
using the fitted linear regression model.

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

```{r}
# Compute RMSE and R-squared for the model using all predictors
all_metrics <- body_predictions %>%
  metrics(truth = BodyWeight, estimate = .pred)

all_rmse <- all_metrics %>%
  filter(.metric == "rmse") %>%
  pull(.estimate)

all_r_squared <- all_metrics %>%
  filter(.metric == "rsq") %>%
  pull(.estimate)
```

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

Finally, we print the computed RMSE and R-squared values to evaluate the
model's performance. These metrics provide insights into how well the
linear regression model fits the data and predicts the outcome variable.

```{r}
# Print RMSE and R-squared for the model using all predictors
cat("Model using all predictors:\n")
cat("RMSE:", all_rmse, "\n")
cat("R-squared:", all_r_squared, "\n")

body_predictions
tidy(body_fit)
```

# Results
The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954. The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy. Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power. Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight. interpretation may be necessary depending on the specific context and goals of the analysis.

### 2024-04-09 products/manuscript/Manuscript.qmd
Project 4

@@
 ---
-title: ""
-subtitle: ""
+title: "Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis"
+subtitle: "Insights into Plasma Lipid Profiles and Immune Cell Dynamics for Cardiovascular Health"
 author: Mutsa Nyamuranga
-date: "02/01/2024"
+date: "`r Sys.Date()`"
 format:
@@
     highlight-style: github
-bibliography: ../assets/dataanalysis-references.bib
-csl: ../assets/apa.csl
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/apa.csl
 ---
 
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
+```{r, echo=FALSE, message=FALSE}
+# load a few R packages
+library(here)
+library(knitr)
+```
 
@@
 ## Description of data and data source
+
 The data I will be using is based on:
@@
 The question I want to answer using the data is whether there is a correlation between circulating β-carotene levels, BCO1 genetic variants, and the progression or resolution of atherosclerosis.
-Specifically, I am interested in examining how these factors relate to plasma lipid profiles (e.g., cholesterol, triglycerides) and immune cell populations, particularly Treg levels, in individuals with atherosclerosis.
+Specifically, I am interested in examining how these factors relate to plasma lipid profiles (e.g., cholesterol, triglycerides).
 The main outcome of interest will be the degree of atherosclerosis progression or resolution, potentially measured by imaging techniques or clinical assessments of plaque burden.
@@
 ## Data aquisition
+
 The data was acquired online from Pinos et.
@@
 
-```{r load_image, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
+```{r load_image_1, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
 library(png)
@@
 
-```{r load_image, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
+```{r load_image_2, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
 library(png)
@@
 ```
+
 Linear Regression: The results reveal the estimated effects of different treatment categories on the response variable, body weight, along with their statistical significance.
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+
+*Summarize what you did, what you found and what it means.*
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+
+*Discuss what you perceive as strengths and limitations of your analysis.*
 
 ## Conclusions
-_What are the main take-home messages?_
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+*What are the main take-home messages?*
+
+*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
 

### 2024-04-09 R/analysis-code/statistical-analysis.R

@@ -1,71 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(here) #for data loading/saving
-library(dplyr)
-library(skimr)
-library(ggplot2)
-library(tidyr)
-library(forcats)
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#Load the data.
-
-#Path to data. Note the use of the here() package and not absolute paths
-figdata <- here::here("data","processed_data","processeddata.rds")
-#load data
-explorfigdata <- readRDS(figdata)
-body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-linCat <- lm(BodyWeight ~ Category, body_weights)  
-
-# place results from fit into a data frame with the tidy function
-linCattable <- broom::tidy(linCat)
-
-#look at fit results
-print(linCattable)
-
-# save fit results table  
-table_linCat = here("results", "resulttable1.rds")
-saveRDS(linCattable, file = table_linCat)
-
-############################
-#### Second model fit
-
-# Assuming your time series data is stored in your_data with Week as the time variable and BodyWeight as the data
-
-# Load required library
-library(strucchange)
-
-# Fit a linear regression model with treatment variables
-timeseries <- lm(BodyWeight ~ Week + I(Week^2) + I(Week^3) + Category, data = body_weights)
-
-# Summary of the model
-summary(timeseries)
-
-# place results from fit into a data frame with the tidy function
-timeseriestbl <- broom::tidy(timeseries)
-
-#look at fit results
-print(timeseriestbl)
-
-# save fit results table  
-seriestable_file2 = here("results", "resulttable2.rds")
-saveRDS(timeseriestbl, file = seriestable_file2)
-
-  

### 2024-04-11 R/eda-code/exploratory-analysis.qmd
Part 4 edits

--- 2024-03-16 R/analysis-code/exploratory-analysis.qmd
@@
   theme_minimal()
+figure_file2 = here("results", "figures", "category_distro.png")
+ggsave(filename = figure_file2, plot=bar1) 
 ```

### 2024-04-11 R/processing-code/processingfile-v1.qmd
Part 4 edits

@@
 ```{r}
-dir.create("/Users/mutsa_n/Desktop/MADA-course/Nyamuranga-MADA-project/data/processed_data/", recursive = TRUE)
 save_data_location <- here::here("data","processed_data","processeddata.rds")

### 2024-04-11 products/manuscript/Manuscript.qmd
Part 4 edits

@@
 library(knitr)
+library(tidyr)
+library(broom)
 ```
@@
 # Load necessary package
-library(readxl)
-
-# Read Excel file
-rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
+# path to data
+# note the use of the here() package and not absolute paths
+data_location <- here::here("data/raw-data/elife-87430-data1-v1.xlsx")
+rawdata <- readxl::read_excel(data_location)
 ```
@@
 
-## Data aquisition
+## Data Acquisition
 
@@
 
-## Data import and cleaning
+## Data Import and Cleaning
 
@@
 ```
+
+### Time Series Regression Analysis
+A time series regression analysis was conducted to investigate temporal trends and relationships between variables over time.
+The time series data were examined for stationarity and seasonality using appropriate statistical tests.
+The primary variables of interest were body weight over weeks.
+A linear regression model was fitted to the time series data to quantify the relationship between the predictor variables and the outcome variable over the study period.
+Model diagnostics, including residual analysis and goodness-of-fit measures, were performed to assess model adequacy.
+
+### Basic Linear Regression Analysis
+A basic linear regression analysis was performed to examine the association between predictor variables and the outcome variable.
+The predictor variables included control, baseline and beta-carotene, while the outcome variable was body weight.
+Prior to analysis, multicollinearity diagnostics were conducted to identify and mitigate issues arising from collinear predictor variables.
+The linear regression model was fitted to the data, and model assumptions were assessed through residual analysis and goodness-of-fit tests.
+
+### Train/Test Linear Regression Analysis
+To assess the generalizability and predictive performance of the linear regression model, a train/test approach was employed.
+The dataset was randomly divided into a training set (80% of the data) and a testing set (remaining 20 of the data).
+The linear regression model was fitted to the training set, and model performance was evaluated using the testing set.
+Performance metrics such as mean squared error (MSE) and R-squared were calculated to assess the predictive accuracy of the model.
+
+### Statistical Software
+All statistical analyses were conducted using [name of statistical software or programming language, e.g., R, Python]. Statistical significance was determined at the conventional alpha level of 0.05.
 
@@
 
+```{r}
+#| label: Full Model Fit
+#| tbl-cap: "Full model fit table."
+#| echo: FALSE
+resulttable3 = readRDS(here("results","output","resulttable3.rds"))
+
+print(resulttable3)
+```
+The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954.
+The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy.
+Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power.
+Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight.
+
+The results highlight the effectiveness of the predictive model in capturing the relationship between predictors and body weight.
+The preprocessing steps, including dummy encoding, centering, and scaling, played a crucial role in enhancing model performance by standardizing and transforming predictors.
+The high explanatory power of the model suggests that the included predictors effectively explain variations in body weight.
+
 {{< pagebreak >}}
@@
 ## Summary and Interpretation
-
-*Summarize what you did, what you found and what it means.*
 

### 2024-04-25 R/eda-code/exploratory-analysis.qmd
Part 5

@@
 
+```{r}
+explorfigdata
+```
+
 # Data exploration through tables
@@
 ```{r}
-summary_explorfigdata = skimr::skim(explorfigdata)
-print(summary_explorfigdata)
+print(explorfigdata)
+
+summary_table <- explorfigdata %>%
+  tbl_summary() %>%
+  add_stat(label = "Count", statistic = list(all_continuous() ~ "{n}")) %>%
+  add_stat(label = "Mean", statistic = list(all_continuous() ~ "{mean}"))
+```
+
 # save to file
 summarytable_elife = here("results", "summarytable.rds")
-saveRDS(summary_explorfigdata, file = summarytable_elife)
+saveRDS(explorfigdata, file = summarytable_elife)
 ```

### 2024-04-25 products/manuscript/Manuscript.qmd
Part 5

@@
 library(here) #to set paths
+library(gtsummary) #for summary tables
+library(tidyverse)
 
@@
 summarytable = readRDS(here("results","tables","summarytable.rds"))
-knitr::kable(summarytable)
+summarytable <-
+  explorfigdata %%
+   tbl_summary(include = c(Week, BaselineF, ControlF, BetaCaroteneF, BaselineM, ControlM, BetaCaroM))
 ```
@@
 ## Summary and Interpretation
+Atherosclerosis, a complex chronic inflammatory condition affecting arterial walls, remains a significant contributor to cardiovascular morbidity and mortality globally.
+Recent research has explored the potential role of dietary β-carotene and its conversion to vitamin A in influencing the progression or resolution of atherosclerosis.
+Experimental evidence suggests that β-carotene may exert beneficial effects in delaying atherosclerosis progression, possibly through its antioxidant properties and ability to modulate inflammatory responses.
+Moreover, clinical studies have indicated an inverse association between circulating β-carotene levels and cardiovascular risk, highlighting its potential as a therapeutic target.
+
+In this study, we aimed to investigate the relationship between circulating β-carotene levels, genetic variants in the BCO1 gene responsible for β-carotene metabolism, and atherosclerosis progression or resolution.
+By analyzing data from a cohort of individuals with varying degrees of atherosclerosis, we observed significant associations between circulating β-carotene levels, certain BCO1 genetic variants, and atherosclerosis severity.
+Regression analyses revealed the predictive value of circulating β-carotene levels and specific BCO1 genetic variants in modulating atherosclerosis, independently of traditional risk factors.
+Additionally, the findings underscored the importance of considering plasma lipid profiles and immune cell populations as potential mediators of β-carotene's effects on atherosclerosis.
+
+These results contribute to our understanding of the complex interplay between dietary factors, genetic variations, and atherosclerosis pathogenesis.
+They suggest that interventions targeting β-carotene metabolism pathways may hold promise for mitigating atherosclerosis progression and reducing cardiovascular risk.
+Further research is warranted to elucidate the mechanistic underpinnings of β-carotene's effects and to validate the findings in larger, longitudinal studies.
 
@@
 
-*Discuss what you perceive as strengths and limitations of your analysis.*
+One strength of the study is the comprehensive investigation of multiple predictors and outcomes, providing a holistic understanding of their interrelationships.
+The use of regression modeling allowed for robust statistical assessment of the associations between circulating β-carotene levels, BCO1 genetic variants, and atherosclerosis progression or resolution.
+
+However, several limitations should be considered.
+Firstly, the study relies on observational data, which may be subject to confounding and bias.
+Additionally, the dataset may lack sufficient granularity or sample size to fully capture the complexities of the relationships under study.
+Furthermore, the cross-sectional nature of the data limits the ability to infer causality, highlighting the need for longitudinal studies to validate the findings.
+Finally, while we accounted for potential confounding variables in the analyses, residual confounding may still exist.
+
+Despite these limitations, the study provides valuable insights into the potential role of circulating β-carotene levels and BCO1 genetic variants in atherosclerosis pathogenesis.
+Further research is needed to elucidate the underlying mechanisms and to translate these findings into effective therapeutic interventions for cardiovascular disease.
 
@@
 
-*What are the main take-home messages?*
-
-*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
+ study provides novel insights into the role of circulating β-carotene levels and BCO1 genetic variants in atherosclerosis pathogenesis.
+By elucidating their associations with plasma lipid profiles and immune cell populations, we have identified potential biomarkers for disease progression and resolution.
+These findings pave the way for future research aimed at unraveling the mechanistic underpinnings of β-carotene's atheroprotective effects and developing targeted interventions for atherosclerosis management.
+Ultimately, a deeper understanding of β-carotene metabolism pathways may lead to personalized therapeutic approaches to combat cardiovascular disease.
 
@@
 # References
+
+Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.
+
+Shaish A, Daugherty A, O'Sullivan F, Schonfeld G, Heinecke JW. Beta-carotene inhibits atherosclerosis in hypercholesterolemic rabbits. J Clin Invest. 1995;96(4):2075-2082. doi:10.1172/JCI118256
+
+Getz GS, Reardon CA. Animal models of atherosclerosis. Arterioscler Thromb Vasc Biol. 2012;32(5):1104-1115. doi:10.1161/ATVBAHA.111.237693
+
+Grune T, Lietz G, Palou A, et al. Beta-carotene is an important vitamin A source for humans. J Nutr. 2010;140(12):2268S-2285S. doi:10.3945/jn.109.119024
+
+Henning RJ.
+Obesity and obesity-induced inflammatory disease contribute to atherosclerosis: a review of the pathophysiology and treatment of obesity.
+Am J Cardiovasc Dis.
+2021;11(4):504-529.
+Published 2021 Aug 15.
+
+Vakhtangadze T, Singh Tak R, Singh U, Baig MS, Bezsonov E.
+Gender Differences in Atherosclerotic Vascular Disease: From Lipids to Clinical Outcomes.
+Front Cardiovasc Med.
+2021;8:707889.
+Published 2021 Jun 28.
+doi:10.3389/fcvm.2021.707889

### 2024-04-25 products/manuscript/supplement/Supplementary-Material.qmd
Part 5

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"

subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis_template_references.bib
csl: ../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

# Overview

A quick overview of what readers can find in the supplement.

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.

### 2024-04-30 R/eda-code/exploratory-analysis.qmd

@@
 library(forcats)
+library(gtsummary) #line added by Rachel for add_stat function to work
 ```
@@
 
+```{r}
 # save to file

### 2024-05-04 R/analysis-code/modeling.qmd
Part 4 edits
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

@@
 library(parsnip)
+library(broom.mixed) # for converting bayesian models to tidy tibbles
+library(dotwhisker)  # for visualizing regression results
+library(jtools) # Load jtools
 ```
@@
 ## First model fit
+
 ```{r}
-# fit linear model using height as outcome, weight as predictor
+# Define the recipe
+body1_recipe <- recipe(BodyWeight ~ Category, data = body_weights) %>%
+  step_dummy(all_nominal(), -all_outcomes()) %>%
+  step_center(all_predictors()) %>%
+  step_scale(all_predictors())
 
-linCat <- lm(BodyWeight ~ Category, body_weights)
+# Split data into training and testing sets
+set.seed(123)  # for reproducibility
+body1_split <- initial_split(body_weights, prop = 0.75, strata = BodyWeight)
+body1_train <- training(body1_split)
+body1_test <- testing(body1_split)
+
+str(body1_test)
+str(body1_train)
+
+# Define the linear regression model specification
+linear_spec <- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+# Create a workflow that incorporates the recipe and the model
+body1_workflow <- workflow() %>%
+  add_recipe(body1_recipe) %>%
+  add_model(linear_spec)
+```
+The fit function fits the workflow to the training data. This step
+applies the preprocessing steps defined in the recipe to the training
+data and then fits the linear regression model to the preprocessed data.
+
+The predict function generates predictions from the fitted model using
+the testing data. This step applies the preprocessing steps (from the
+recipe) to the testing data and then predicts the outcome variable (Y)
+using the fitted linear regression model.
+
+The metrics function computes evaluation metrics for the model
+predictions. Here, we calculate the root mean squared error (RMSE) and
+the R-squared value to assess the model's performance. The computed
+metrics are then extracted and stored in the variables all_rmse and
+all_r_squared, respectively.
+
+```{r}
+# Convert BodyWeight to numeric if it's stored as character
+body1_train$BodyWeight <- as.numeric(body1_train$BodyWeight)
+
+# Convert BodyWeight to numeric if it's stored as character
+body1_test$BodyWeight <- as.numeric(body1_test$BodyWeight)
+
+# Fit the workflow to the training data
+linCat <- fit(body1_workflow, data = body1_train)
+
+# Compute predictions on the testing data
+body1_preds <- predict(linCat, new_data = body1_test) %>%
+  bind_cols(body1_test)
+```
+
+```{r}
+# Compute RMSE and R-squared for the model
+all_metrics <- body1_preds %>%
+  metrics(truth = BodyWeight, estimate = .pred)
+
+all_rmse <- all_metrics %>%
+  filter(.metric == "rmse") %>%
+  pull(.estimate)
+
+all_r_squared <- all_metrics %>%
+  filter(.metric == "rsq") %>%
+  pull(.estimate)
+```
+
+```{r}
+# Print RMSE and R-squared for the model
+cat("Model:\n")
+cat("RMSE:", all_rmse, "\n")
+cat("R-squared:", all_r_squared, "\n")
+
+all_rmse
+all_r_squared
+tidy(linCat)
+```
+The RMSE is 5.77076
+
+### Model Tuning
+So, I will tune the model using cross-validation
+
+The model performance is evaluated using the
+cross-validation technique employing 10 folds for the two main models.
+This technique involves dividing the training data into 10 subsets and
+iteratively using 9 of them for model fitting while reserving the
+remaining 10% for evaluation. This process is repeated 10 times,
+ensuring that each subset serves as both training and evaluation data.
+
+```{r}
+# Train the model using train function from caret package
+set.seed(234)  # for reproducibility
+folds1 <-vfold_cv(body1_train, v=10)
+folds1
+```
+Subsequently, an object for resampling is constructed using the
+'workflow' function from tidymodels. This function combines
+pre-processing, modeling, and post-processing instructions into a single
+entity, streamlining the analysis pipeline.
+
+```{r}
+#Resampling using workflow for the model with only DOSE as predictor
+linCat2 <- 
+	workflow() %>%
+	add_model(linear_spec) %>%
+  add_formula(BodyWeight ~ Category)%>%
+	fit_resamples(folds1)
+
+collect_metrics(linCat2)
+```
+Using cross-validation, the RMSE increases to 6.0090840.
+
+```{r}
+#| message: false
+#| warning: false
+# Creating a data-frame with observed and predicted values from the model with `Category` as the predictor
+lin1_fit <- linCat %>%
+  predict(body1_train) %>%
+  bind_cols(body1_train)
+
+#I added a label column and dubbed it 'Simple Linear model'
+lin1_fit$label <- rep("Simple Linear Model")
+
+# Create the ggplot figure to graph the predictive values vs the observed value for the three models
+p1 <- ggplot(
+  lin1_fit, aes(x = BodyWeight, y = .pred, color = label, shape = label)) +
+  geom_point(size=2) +
+   scale_color_manual(values = c("#9467bd", "#ff9896", "#17becf"))+
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +  # Adding a 45-degree line
+  labs(x = "Observed Values", y = "Predicted Values", color = "Model", shape = "Model") +
+  xlim(0, 50) + 
+  ylim(0, 50)+
+  theme_bw()
+
+# Viewing the plot
+p1
+```
+Based on the results of the initial model fitting generating a RMSE of 5.77076 and the cross-validation generating a RMSE increase to 6.0090840, we can conclude that this first model has over-fit the data.
+
+With this, I will move on to other models to find a RMSE that works best for assessing the data. 
+```{r}
+figure_file3 = here("results", "figures", "SimpleLinearPlot.png")
+ggsave(filename = figure_file3, plot= p1) 
+```
+
+```{r}
+# Summary of the model
+summary(linCat)
 
 # place results from fit into a data frame with the tidy function
-linCattable <- broom::tidy(linCat)
+fullmodel1 <- broom::tidy(linCat)
 
 #look at fit results
-print(linCattable)
+print(fullmodel1)
 
 # save fit results table  
-table_linCat = here("results", "resulttable1.rds")
-saveRDS(linCattable, file = table_linCat)
+model1_file = here("results", "output", "resulttable1.rds")
+saveRDS(fullmodel1, file = model1_file)
 ```
@@
 saveRDS(timeseriestbl, file = seriestable_file2)
+
+# Convert the results into a data frame
+results <- data.frame(
+  term = c("(Intercept)", "Week", "I(Week^2)", "I(Week^3)", 
+           "CategoryBaselineM", "CategoryBetaCaroM", 
+           "CategoryBetaCaroteneF", "CategoryControlF", "CategoryControlM"),
+  estimate = c(14.963391564, 1.104685173, 0.035042256, -0.002315347,
+               8.629167500, 8.633072223, -1.569384093, -1.044471462, 8.186054854),
+  std.error = c(0.6518010412, 0.2519033316, 0.0292788770, 0.0009802486,
+                0.4469748457, 0.4325930512, 0.4325930512, 0.4325930512, 0.4325930512)
+)
+
+# Calculate the confidence intervals
+results$lower_ci <- results$estimate - 1.96 * results$std.error
+results$upper_ci <- results$estimate + 1.96 * results$std.error
+
+# Plot a forest plot
+library(ggplot2)
+
+ggplot(results, aes(x = estimate, y = term)) +
+  geom_point() +
+  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci)) +
+  labs(title = "Forest Plot of Regression Coefficients",
+       x = "Estimate",
+       y = "Term") +
+  theme_minimal()
 ```
+```{r}
+# Convert BodyWeight to numeric
+body_weights$BodyWeight <- as.numeric(body_weights$BodyWeight)
+
+# Remove observations with missing values
+complete_data <- body_weights[!is.na(body_weights$BodyWeight),]
+
+# Obtain predicted values for complete data
+predicted_time_complete <- predict(timeseries, newdata = complete_data)
+
+# Calculate residuals
+residuals_complete <- complete_data$BodyWeight - predicted_time_complete
+
+# Calculate RMSE
+rmse_complete <- sqrt(mean(residuals_complete^2))
+
+# Print RMSE
+print(rmse_complete)
+```
+The RMSE for this model is 1.210414.
+
+This shows a high level of accuracy in predicting values, however, we can assume that this model may also be over fitting due to an increase level of nuance within the data being taken in to consideration.
+
 ## Third Model Fit
@@
 ```{r}
-body_weights$BodyWeight <- as.numeric(as.character(body_weights$BodyWeight))
+body_train$BodyWeight <- as.numeric(as.character(body_train$BodyWeight))
 
-class(body_weights$BodyWeight)
+class(body_train$BodyWeight)
 ```
@@
   bind_cols(body_test)
+
+body_predictions$BodyWeight <- as.numeric(as.character(body_predictions$BodyWeight))
 ```
@@
 
+```{r}
+# Summary of the model
+summary(body_fit)
+
+# place results from fit into a data frame with the tidy function
+fullmodel <- broom::tidy(body_fit)
+
+#look at fit results
+print(fullmodel)
+
+# save fit results table  
+model2_file = here("results", "output", "resulttable3.rds")
+saveRDS(fullmodel, file = model2_file)
+```
+
 # Results

### 2024-05-04 R/eda-code/exploratory-analysis.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

@@
 ```{r}
-#Path to data. Note the use of the here() package and not absolute paths
+#Path to data.
 figdata <- here::here("data","processed_data","processeddata.rds")
@@
 explorfigdata <- readRDS(figdata)
-```
-
-```{r}
-explorfigdata
 ```
@@
 ```{r}
-print(explorfigdata)
+# Calculate count and mean statistics
+explorfigdata <- mutate_all(explorfigdata, as.numeric)
+# Calculate count statistics
+summary_stats <- explorfigdata %>%
+  summarise(across(starts_with(c("Baseline", "Control", "BetaCaro")), list(Count = ~sum(!is.na(.))), .names = "{.col}_{.fn}"))
 
-summary_table <- explorfigdata %>%
-  tbl_summary() %>%
-  add_stat(label = "Count", statistic = list(all_continuous() ~ "{n}")) %>%
-  add_stat(label = "Mean", statistic = list(all_continuous() ~ "{mean}"))
+# Reshape the data for better presentation
+summary_table <- summary_stats %>%
+  pivot_longer(everything(), names_to = c(".value", "Variable"), names_sep = "_")
+# Print the summary table
+print(summary_table)
+```
+
+This table displays the counts of data per week. Where 16 would represent the number of weeks were data was collected for the Baseline variable mice.
+
+```{r}
+# save to file
+summary_table = here("results", "summarytable1.rds")
+saveRDS(explorfigdata, file = summary_table)
 ```
@@
 ```{r}
+body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
+
+summary_table2 <-
+  tbl_summary(
+    body_weights,
+    by = Category,
+    missing = "no" # don't list missing data separately
+  ) %>%
+  add_n() %>% # add column with total number of non-missing observations
+  modify_header(label = "**Variable**") %>% # update the column header
+  bold_labels()
+
+summary_table2
+
 # save to file
 summarytable_elife = here("results", "summarytable.rds")
-saveRDS(explorfigdata, file = summarytable_elife)
+saveRDS(summary_table2, file = summarytable_elife)
 ```
 
-We are saving the results to the `results` folder.
-Depending on how many tables/figures you have, it might make sense to have separate folders for each.
-And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
-Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
+With some final exploration.
+We look at the change in body weight over the 16 weeks.
+We see that males start and finish with higher body weights as expected.
+Of note, we see that the control and Beta-Carotene condition for female mice ends with the lowest body weights.
 
@@
 # Saving the plot as an image file
-figure_file = here("results","condition_distro.png")
+figure_file = here("results","condition_distro1.png")
 ggsave(filename = figure_file, plot=bar1) 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 # Reshape data to long format
@@
 library(ggplot2)
-ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
+line1 <- ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
   geom_line() +
@@
   theme_minimal()
+
+plot(line1)
+
 figure_file2 = here("results", "figures", "category_distro.png")
-ggsave(filename = figure_file2, plot=bar1) 
+ggsave(filename = figure_file2, plot= line1) 
 ```
-
-With some final exploration.
-We look at the change in body weight over the 16 weeks.
-We see that males start and finish with higher body weights as expected.
-Of note, we see that the control and Beta-Carotene condition for female mice ends with the lowest body weights.
+The line graph of total body weight over time reflects changes in body weight across different time points.
+Even if the sum of baseline body weights is lower, the trend over time may show an increase or decrease in body weight that is not evident when only considering the sum of baseline weights.

### 2024-05-04 R/processing-code/processingfile-v1.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

@@
 # Remove specified columns and rows
-figure1a<- figure1a[, -c(3, 5, 7, 9, 11, 13)]
+figure1a <- figure1a[, -c(3, 5, 7, 9, 11, 13)]
 figure1afixed <- figure1a[-1, ]
+
+# Remove missing data
+figure1afixed <- na.omit(figure1afixed)
 
@@
 
+```{R}
+# Assign the cleaned data frame to a final variable
+cleaned_data <- figdata
+```
+
 # Save data
@@
 
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).
+The missing data was included in order to keep whole categories from being removed. While it may weaken extrapolation in analysis coding, it is necessary for the cleaning process.

### 2024-05-04 products/manuscript/Manuscript.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

@@
 library(broom)
+library(kableExtra)
+library(sjmisc)
+library(sjlabelled)
+library(jtools)
 ```
@@
 
-{{< pagebreak >}}
+This project investigates the intricate interplay between circulating β-carotene levels, genetic variants in the BCO1 gene, and the progression or resolution of atherosclerosis.
+Atherosclerosis, a major contributor to cardiovascular morbidity and mortality, is characterized by the accumulation of cholesterol-rich lipoproteins in arterial walls.
+Epidemiological evidence suggests a potential role for β-carotene in modulating atherosclerosis progression, attributed to its antioxidant properties and immune-modulating effects.
+Leveraging data from a murine model, this study employs regression modeling to assess the predictive value of circulating β-carotene levels and BCO1 genetic variants on atherosclerosis severity.
+Results highlight significant associations between these factors and plasma lipid profiles, shedding light on potential therapeutic targets for cardiovascular disease management.
+Using data from an animal trial, we analyzed plasma lipid profiles, immune cell dynamics, and atherosclerosis severity.
+Regression analyses revealed significant associations between β-carotene levels, BCO1 variants, and atherosclerosis severity, independent of traditional risk factors.
+Findings suggest potential biomarkers for disease progression and resolution, paving the way for targeted interventions in cardiovascular disease management.
+Despite limitations inherent to observational data, this study provides valuable insights into the complex mechanisms underlying atherosclerosis pathogenesis and underscores the need for further research to validate these findings in clinical settings.
 
@@
 
-Atherosclerosis, a chronic inflammatory disease of the arterial wall, remains a leading cause of cardiovascular morbidity and mortality worldwide.
-Despite significant advancements in understanding its pathophysiology, effective therapeutic interventions are still in demand.
-Recent research has shed light on the potential role of dietary β-carotene and its conversion to vitamin A in modulating atherosclerosis progression and resolution.
+Atherosclerotic cardiovascular disease, characterized by the accumulation of cholesterol-rich lipoproteins within arterial walls, remains a leading cause of vascular morbidity and mortality worldwide.
+Its major clinical manifestations include ischemic heart disease, ischemic stroke, and peripheral arterial disease (Herrington et al., 2016).
 
-Several studies have demonstrated the beneficial effects of β-carotene in delaying atherosclerosis progression in experimental models, along with its association with reduced cardiovascular risk in humans.
-Moreover, genetic variants in the BCO1 gene, responsible for converting β-carotene to vitamin A, have been linked to alterations in circulating carotenoid levels and plasma lipid profiles in individuals.
-However, the precise mechanisms underlying these effects, particularly regarding the interplay between β-carotene, BCO1 activity, and atherosclerosis progression or resolution, remain unclear.
+Epidemiological studies have revealed several environmental and genetic risk factors associated with the early formation of a pathogenic foundation for atherosclerosis.
+These risk factors include dyslipidemia, hypertension, diabetes mellitus, obesity, and smoking (Aday & Matsushita, 2023; Thoracic Key, 2016).
+
+In high-income countries, there have been dramatic declines in the incidence and mortality from ischemic heart disease and ischemic stroke since the middle of the 20th century.
+For example, in the United Kingdom, the probability of death from vascular disease in middle-aged men (35–69 years) has decreased from 22% in 1950 to 6% in 2010.
+Most low- and middle-income countries have also reported declines in mortality from stroke over the last few decades, but mortality trends from ischemic heart disease have been more varied, with some countries reporting declines and others reporting increases (particularly those in Eastern Europe and Asia).
+Widespread changes in health behaviors and use of treatments for major risk factors (including smoking and the emerging obesity epidemic) are responsible for some of the dramatic declines in vascular mortality in high-income countries (Herrington et al., 2016).
+
+### The Role of β-Carotene
+β-Carotene, a provitamin A compound found in colorful fruits and vegetables, has attracted attention due to its potential impact on atherosclerosis.
+Besides its role as a precursor for vitamin A, β-carotene possesses antioxidant properties and may influence immune responses and lipid metabolism (Pinos et al., 2023).
+
+### BCO1: The Clever Enzyme
+β-Carotene oxygenase 1 (BCO1) is the enzyme responsible for cleaving β-carotene to form vitamin A.
+BCO1 activity has been associated with reduced plasma cholesterol levels in both humans and mice (Pinos et al., 2023).
+Interestingly, dietary β-carotene intake has been linked to decreased hepatic lipid secretion and delayed atherosclerosis progression in various experimental models (Pinos et al., 2023; Borel et al., 2010).
+
+### Accelerating Atherosclerosis Resolution
+Recent research has revealed an exciting twist: β-carotene not only delays atherosclerosis progression but also accelerates its resolution.
+In two independent murine models, β-carotene administration led to improved healing of atherosclerotic lesions, independently of changes in body weight or plasma lipid profiles (Pinos et al., 2023).
+
+### Vitamin A Production Matters
+To understand this phenomenon, researchers investigated Bco1-deficient mice (Bco1-/-).
+These mice lack the ability to convert β-carotene to vitamin A.
+Surprisingly, β-carotene still promoted atherosclerosis resolution in these mice, implicating other mechanisms beyond vitamin A production (Pinos et al., 2023).
+
+### Regulatory T Cells (Tregs) Enter the Scene
+To explore further, the researchers focused on regulatory T cells (Tregs), known for their role in immune regulation.
+β-Carotene favored Treg expansion within atherosclerotic plaques.
+Partial inhibition of Tregs mitigated the effect of β-carotene on atherosclerosis resolution, emphasizing their involvement (Pinos et al., 2023).
 
@@
 
-The data I will be using is based on:
+The data utilized in this study originates from an animal trial conducted with mice, adhering to approved procedures by the Institutional Animal Care and Use Committees of the University of Illinois at Urbana Champaign.
+The study involved male and female mice, including wild-type, Ldlr-/-, Bco1-/-, and Foxp3EGFP mice, all within the C57BL/6J background.
+Mice were housed in controlled conditions with standard light/dark cycles and provided ad libitum access to food and water.
 
-Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.
+Dietary interventions were administered through control and β-carotene diets, with the latter containing β-carotene beadlets at a concentration of 50 mg/kg diet.
+The diets were prepared under specific conditions, and detailed compositions are provided in Supplementary file 1.
+Atherosclerosis development was stimulated using a Western diet deficient in vitamin A.
 
-The data consists of measurements of circulating β-carotene and retinoic acid levels, along with relevant biomarkers and clinical characteristics, collected from a cohort of individuals with varying degrees of atherosclerosis progression and resolution.
-Genetic information on BCO1 activity, including genetic variants associated with BCO1 activity, may also be included in the dataset.
-The data was collected through blood sample analysis and genetic testing.
+Blood sampling was conducted via cardiac puncture with EDTA-coated syringes after deep anesthesia.
+Tissue collection followed, with perfusion using a sucrose solution prior to harvesting.
+Tissues were snap-frozen in liquid nitrogen and stored at -80°C.
 
-```{r}
+The data extracted for analysis was from a table in the first sheet of a supplementary Excel file associated with the study.
+Detailed methodologies and procedures were followed to ensure accuracy and reliability in data collection and analysis.
+
+```{r, echo=FALSE, message=FALSE}
 # Load necessary packages
@@
 
-{{< pagebreak >}}
-
 # Methods
@@
 
-```{r load_image_1, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
-library(png)
-condimage <- readPNG(here("results", "figures", "condition_distro.png"))
-par(mar = c(1, 1, 1, 1))
-plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
-rasterImage(condimage, 0.1, 0.1, 0.9, 0.9)  
-```
-
-```{r load_image_2, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
-library(png)
-catimage <- readPNG(here("results", "figures", "category_distro.png"))
-par(mar = c(1, 1, 1, 1))
-plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
-rasterImage(catimage, 0.1, 0.1, 0.9, 0.9)  
-```
-
-### Time Series Regression Analysis
-A time series regression analysis was conducted to investigate temporal trends and relationships between variables over time.
-The time series data were examined for stationarity and seasonality using appropriate statistical tests.
-The primary variables of interest were body weight over weeks.
-A linear regression model was fitted to the time series data to quantify the relationship between the predictor variables and the outcome variable over the study period.
-Model diagnostics, including residual analysis and goodness-of-fit measures, were performed to assess model adequacy.
-
-### Basic Linear Regression Analysis
-A basic linear regression analysis was performed to examine the association between predictor variables and the outcome variable.
-The predictor variables included control, baseline and beta-carotene, while the outcome variable was body weight.
-Prior to analysis, multicollinearity diagnostics were conducted to identify and mitigate issues arising from collinear predictor variables.
-The linear regression model was fitted to the data, and model assumptions were assessed through residual analysis and goodness-of-fit tests.
-
-### Train/Test Linear Regression Analysis
-To assess the generalizability and predictive performance of the linear regression model, a train/test approach was employed.
-The dataset was randomly divided into a training set (80% of the data) and a testing set (remaining 20 of the data).
-The linear regression model was fitted to the training set, and model performance was evaluated using the testing set.
-Performance metrics such as mean squared error (MSE) and R-squared were calculated to assess the predictive accuracy of the model.
+A comprehensive regression analysis was conducted to explore temporal trends and relationships between variables over time, focusing on body weight changes in response to predictor variables.
+Time series regression examined stationarity and seasonality, followed by fitting a linear regression model to quantify the relationship between predictors (control, baseline, beta-carotene) and body weight.
+Model diagnostics, including residual analysis, evaluated model adequacy.
+Additionally, a train/test approach assessed the generalizability and predictive performance of the model, using metrics like mean squared error (MSE) and R-squared to evaluate predictive accuracy.
 
@@
 All statistical analyses were conducted using [name of statistical software or programming language, e.g., R, Python]. Statistical significance was determined at the conventional alpha level of 0.05.
-
-{{< pagebreak >}}
 
@@
 summarytable = readRDS(here("results","tables","summarytable.rds"))
-summarytable <-
-  explorfigdata %%
-   tbl_summary(include = c(Week, BaselineF, ControlF, BetaCaroteneF, BaselineM, ControlM, BetaCaroM))
+knitr::kable(summarytable)
 ```
+This is data summary table show the distribution of variables by treatment category. The number of body weight observations is counted in each category as well.
 
-## Basic statistical analysis
+```{r load_image_1, echo=FALSE, fig.cap="Condition Distribution" , fig.width=8, fig.height=6 }
+library(png)
+condimage <- readPNG(here("results", "figures", "condition_distro1.png"))
+par(mar = c(1, 1, 1, 1))
+plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
+rasterImage(condimage, 0.1, 0.1, 0.9, 0.9)  
+```
+This graph shows the total body weights in each category with the Male conditions highlighted blue and the female condition highlighted red.
+Here, we see that the beta-carotene condition for males has the highest overall total body weight, while the baseline female condition has the lowest weight.
+```{r load_image_2, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
+library(png)
+catimage <- readPNG(here("results", "figures", "category_distro.png"))
+par(mar = c(1, 1, 1, 1))
+plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
+rasterImage(catimage, 0.1, 0.1, 0.9, 0.9)  
+```
+This graph demonstrated the rate of change in body weight over time in weeks.
+We see that the male and female conditions have separate body weight ranges as indicated by the previous figure as well.
+Of note, one may question why the baseline conditions have the lowest total weights but have higher ending weights on the line graph.
+Even if the sum of baseline body weights is lower, the trend over time may show an increase or decrease in body weight that is not evident when only considering the sum of baseline weights.
 
-## Full analysis
+## Statistical analysis
 
-```{r}
+### Linear Regression
+A basic linear regression analysis was performed to examine the association between predictor variables and the outcome variable.
+The predictor variables included control, baseline and beta-carotene, while the outcome variable was body weight.
+Prior to analysis, multicollinearity diagnostics were conducted to identify and mitigate issues arising from collinear predictor variables.
+The linear regression model was fitted to the data, and model assumptions were assessed through residual analysis and goodness-of-fit tests.
+
+```{r, message=FALSE,warning=FALSE}
 #| label: linreg
-#| tbl-cap: "Linear model fit table."
+#| tbl-cap: "Simple Linear Model."
 #| echo: FALSE
 resulttable1 = readRDS(here("results","output","resulttable1.rds"))
+
 knitr::kable(resulttable1)
@@
 
-Linear Regression: The results reveal the estimated effects of different treatment categories on the response variable, body weight, along with their statistical significance.
+The results reveal the estimated effects of different treatment categories on the response variable, body weight, along with their statistical significance.
 The intercept, estimated at 25.0, represents the baseline level of the response variable when all predictors are zero.
@@
 Conversely, "BetaCaroteneF", "ControlF", and "ControlM" exhibit smaller or even negative effects, but the statistical significance varies among these categories, with "BetaCaroteneF" showing a non-significant effect and others demonstrating statistically significant effects with low p-values.
+
+The initial model demonstrates unfavorable performance, as indicated by an RMSE of 5.77076 and a subsequent increase to 6.0090840 in the cross-validation phase.
+These metrics suggest that the model has over-fit the data, meaning it performs well on the training dataset but poorly on new, unseen data.
+The high RMSE values indicate substantial deviation between predicted and actual body weight values, signaling poor predictive accuracy.
+Additionally, over-fitting hampers the model's generalizability and its ability to capture the underlying patterns in the data, leading to unreliable predictions.
+In contrast to the favorable performance metrics of the predictive model mentioned earlier, these results underscore the limitations of the simple linear model and highlight the need for more robust modeling techniques to accurately predict body weight.
+
+### Time Series Regression Analysis
+A time series regression analysis was conducted to investigate temporal trends and relationships between variables over time.
+The time series data were examined for stationarity and seasonality using appropriate statistical tests.
+The primary variables of interest were body weight over weeks.
+A linear regression model was fitted to the time series data to quantify the relationship between the predictor variables and the outcome variable over the study period.
+Model diagnostics, including residual analysis and goodness-of-fit measures, were performed to assess model adequacy.
 
@@
 
-Time Series Model: The linear regression model was constructed to investigate the relationship between body weight and time (Week) while considering treatment variables.
+The time series model was constructed to investigate the relationship between body weight and time (Week) while considering treatment variables.
 The model revealed several key findings.
@@
 
-```{r}
+The Root Mean Square Error (RMSE) of 1.210414 indicates that, on average, the model's predictions deviate from the actual values by approximately 1.21 units.
+Typically, a lower RMSE signifies better accuracy, so in this case, the RMSE suggests a relatively high level of accuracy in predicting values.
+
+However, the concern of overfitting arises due to the possibility that the model might be capturing noise or random fluctuations in the training data rather than the underlying patterns.
+Overfitting occurs when a model learns the training data too well, to the extent that it performs poorly on unseen data or fails to generalize.
+
+### Full Linear Regression Analysis
+To assess the generalizability and predictive performance of the linear regression model, a train/test approach was employed.
+The dataset was randomly divided into a training set (75% of the data) and a testing set (remaining 25% of the data).
+The linear regression model was fitted to the training set, and model performance was evaluated using the testing set.
+Performance metrics such as mean squared error (MSE) and R-squared were calculated to assess the predictive accuracy of the model.
+
+```{r, message=FALSE,warning=FALSE}
 #| label: Full Model Fit
@@
 
-print(resulttable3)
+knitr::kable(resulttable3)
 ```
@@
 The high explanatory power of the model suggests that the included predictors effectively explain variations in body weight.
-
-{{< pagebreak >}}
 
@@
 Additionally, the findings underscored the importance of considering plasma lipid profiles and immune cell populations as potential mediators of β-carotene's effects on atherosclerosis.
+
+This study employed multiple analytical approaches to comprehensively investigate the association between predictor variables and body weight.
+While each analysis offered unique insights, collectively, they underscored the complex interplay of factors influencing body weight variations, including treatment interventions and temporal trends.
+Despite the limitations of individual analyses, the findings provide valuable implications for managing and predicting body weight changes, highlighting the importance of robust modeling techniques and comprehensive data analysis in understanding complex phenomena like body weight dynamics.
 
@@
 doi:10.3389/fcvm.2021.707889
+
+Luca, A., David, S.
+G., David, A.
+G., Ţarcă, V., Pădureț, I., Mîndru, D.
+E., Roșu, S.
+T., Roșu, E.
+V., Adumitrăchioaiei, H., Bernic, J., Cojocaru, E., & Ţarcă, E.
+(2023).
+Atherosclerosis from Newborn to Adult—Epidemiology, Pathological Aspects, and Risk Factors.
+Life, 13(10), 2056.
+https://doi.org/10.3390/life13102056
+
+Themes, U. (2016, May 27). Atherosclerosis: Epidemiology and Pathophysiology. Thoracic Key. https://thoracickey.com/atherosclerosis-epidemiology-and-pathophysiology/
+
+Herrington, W.
+G., Lacey, B., Sherliker, P., Armitage, J., & Lewington, S.
+(2016).
+Epidemiology of atherosclerosis and the potential to reduce the global burden of atherothrombotic disease.
+Circulation Research, 118(4), 535–546.
+https://doi.org/10.1161/circresaha.115.307611

### 2024-05-04 reviews/project-review-Robertson.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

---
title: "Rachel Robertson Project Review"
author: "Rachel Robertson"
date: 4-26-24
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: "Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis"

Name of project author(s): Mutsa Nyamuranga

Name of project reviewer: Rachel Robertson

# Specific project content evaluation

## Background, Context and Motivation
>How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

I think the background of the project is described in a clear and concise manner, but you may also consider adding statistics about atherosclerosis, and its prevalence, to strengthen your rationale for your study.

The previous study, from which the data came from, could be elaborated on a bit more. Why did they conduct this study with mice? Is this an initial test for drug candidacy or has beta-carotene made it into human clinical trials?

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* some contextualization and motivation

## Question description
>How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

I understand your research question of addressing how these factors are correlated with atherosclerosis progression, but I am curious about what your rationale for choosing this question is. Is this correlation yet to be addressed in literature? Is there applicability to future clinical trials?

I would also suggest rewording the last few sentences of your Questions/Hypothesis section to sound more absolute in the method of analysis you will use (or instead just talk about analysis method in the methods).

### Summary assessment

* question/hypotheses fully clear

## Data description
>How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

I understand that the data came from an animal trial, using mice, but I do not understand much other context about the data. However, I do see that the source is mentioned. I would recommend going into more detail about how the animal trials were conducted and how the meaasurements for each indicator were collected.

I do not see a codebook for your variables of interest. Does your data source include a codebook or will you include a codebook in your readme file under the R folder?

### Summary assessment

* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
>How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

I think that some of the cleaning steps could use additional rationale in the code file. For example, what was your rationale for not getting rid of missing values? I know sometimes it's not the best, but for your data, is there a reason to keep missing values?

I also saw an interesting trend in your EDA. What might be the reason that the sum of body weights for baseline males and females is the lowest, but they are not the lowest in the line graph of total body weight over time?

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
>Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

I think you can be a little bit more specific with what effects you are talking about for each category, in the linear fit results section.For example, it would make it more clear to list out each variable and explain their effect on the response, noting the variables with the larger p-values (but without invalidating them, because they likely still have importance).

What is the rationale behind your choice of analysis method? Why did you choose to use a linear regression and a time series regression to analyze body weight? Were there not many outlines and a normal distribution for the variables of interest? How did the linear model and time series models perform compared to one another?

### Summary assessment

* defensible but not optimal analysis

## Presentation
>How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

One of the EDA figures, total weight for conditions, is displayed twice in your manuscript. I think you may want to delete one of these. You may also consider changing the colors or font of this graph. I would also recommend rearranging the order of the factors in the legend part of the graph, so all male and female variables are grouped together.

I also think it might be useful to get rid of some of the code in the manuscript, and replace it with just the summary tables or specific output (in table or figure form), that you want to display.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions
>Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion has not been edited, but the final paragraph of the results sounds like the beginnings of a discussion. You might want to move this and continue to elaborate upon it in the discussion section.

### Summary assessment

* major parts of discussion missing or wrong

## Further comments

You might want to consider trying another analysis method to examine whether it performs better. For example, you might see if a GAm gives you similar results to the linear regression.Although, the main thing you should focus on is finishing and fleshing out the manuscript.

# Overall project content evaluation

## Structure
>Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is organized into folders, but I found the readme files to be unhelpful. The main README file should be updated to explain what is in all of your folders and in what order to run the code. The readme in raw-data, the readme(s) in the main R folder and all of the analysis sub-folders, and in the readme in the results folder and sub-folders should be updated to reflect what is contained in them.

The results folder is a bit unorganized, with tables and figures in the main folder, as well as sub-folders. Make sure that your save paths are correct and delete the files saved int he incorrect place.

Under the products/ manuscript folder, the supplementary material may be updated.

### Summary assessment

* poor and confusing structure

## Documentation 
>How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Most of the code was well explained and documented with comments in the chunks as well as between chunks. There were, however, a few exceptions:
Some chunks of code in EDA and processing do not specify the purpose of the code. I think it would be helpful to add explanations in your lines of code with a "#" to describe what the functions do.

### Summary assessment

* decently documented with some gaps

## Reproducibility
>Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The cleaning code worked fine for me, but the exploratory data analysis did not. First I had to open the gtsummary library to get the add_stat function, but after this it said you are missing an argument (fns=) in the add_stat function. I was not able to add the argument and fix it, but it should be specified according to type of statistic you are looking to add to your table. I also had trouble rendering your manuscript because it was not able to pull the summary table (because of the previous error in the eda code the summary table data was not produced). Instead, I opened your manuscript docx.

The rest of the EDA and the analysis code worked well and was easily reproducible for me!

### Summary assessment

* small parts not reproducible or required manual intervention

## Thoroughness
>How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

I do see a linear regression model used with TidyModel considered, but no further models.Was the time series model produced in the eda-code file?

Two models were considered, but the rationale for considering these is not sufficiently explained. It would be useful to look at alternative ways to model the data to find the most important predictor of your outcome. I like using the vip package for finding the most important variables using the ml models. I would consider looking into this and also testing out another similar model that makes sense with your data, like a RF regression model or a GAM.

### Summary assessment

* weak level of thoroughness

### 2024-05-04 reviews/project-review-by CLi.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

---
title: Project Review Part 5 
author: Chaohua Li
date: 04-25-2024
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis

Name of project author(s): Mutsa Nyamuranga

Name of project reviewer: Chaohua Li

# Instructions

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background needs to be more detailed. It would also be helpful with some citations added.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* some contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The research question was clearly stated and explained.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Data source was clearly described. There was no codebook provided.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data was well processed. Detailed comments/notes were provided for all procedures.
There was no supplementary material. The supplement qmd file in the 'manuscript' folder was from the template file.

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis methods were appropriate for the question to be answered. Cross-validation was not used for performance evaluation. But the performance of the final was assessed on a test dataset.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results were presented in a well-structured manner. Most tables were well-formated. Maybe Table 1(summary of data) should be made more concise or moved into a supplemental material.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Discussion is detailed and the interpretation of results is appropriate. Strengths and limitations are specified.

### Summary assessment

* strong, complete and clear discussion

## Further comments

There are some R scripts showing up in the manuscript. And some residual content of the template needs to be removed. It would be helpful if the detailed titles/annotations can be added to the figures.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is generally well-structured. All files are in well labeled folders. It is easy to figure out the use for each file/folder.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The R scripts are well-documented with sufficient description. Each decision of analysis was justified and explained. There are residual content from template in many Rmd files(README). The qmd file for supplement material is also from the template.

### Summary assessment

* decently documented with some gaps

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was able reproduce the results. Instructions were clear.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The research questions have been fully addressed. All sections are relatively complete and thorough.

### Summary assessment

* strong level of thorougness

## Further comments

Overall, this study is in good shape. The author may want to consider moving some content into a supplement file. Some R scripts show up in the manuscript, so that needs to be fixed. The residual content from template in many Rmd files also need to be cleaned.

### 2024-05-04 reviews/project-review-kaggwa_patrick_musta.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

--- 2024-05-04 reviews/project-review-by CLi.qmd
@@
 ---
-title: Project Review Part 5 
-author: Chaohua Li
-date: 04-25-2024
+title: Project Review Template 
+author: Patrick Kaggwa
+date: 4-25-2024
 format: 
@@
 
-Title of project: Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis
+Title of project:Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis (Insights into Plasma Lipid Profiles and Immune Cell Dynamics for Cardiovascular Health)
 
@@
 
-Name of project reviewer: Chaohua Li
+Name of project reviewer: Patrick Kaggwa
 
 # Instructions
+
+Write your comments and feedback below for each section/component of the project.
+The goal should be to help the author improve their project.
+Make comments as constructive and actionable as possible.
+You can provide both criticism and praise.
+
+For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.
+
+Make sure your final document compiles/renders into a readable, well-formatted html document.
+
+Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)
 
@@
 
-The background needs to be more detailed. It would also be helpful with some citations added.
+Overall, background demonstrates a strong contextualization of the research within the existing literature and a clear motivation for the study. Though the text is missing some references.
 
@@
 
-* some contextualization and motivation
+* strong contextualization and motivation
 
@@
 
-The research question was clearly stated and explained.
+Overall, the questions and hypotheses are well-defined and directly relate to the data, providing a clear framework for the analysis and interpretation of results
 
@@
 
-Data source was clearly described. There was no codebook provided.
+The data is adequately described, providing essential information about its source and structure.
 
@@
 
-* source and overall structure of data well explained
+* source and overall structure of data somewhat explained
 
@@
 
-Data was well processed. Detailed comments/notes were provided for all procedures.
-There was no supplementary material. The supplement qmd file in the 'manuscript' folder was from the template file.
+Overall, the data cleaning and processing steps are reasonable.
 
@@
 
-The analysis methods were appropriate for the question to be answered. Cross-validation was not used for performance evaluation. But the performance of the final was assessed on a test dataset.
+Overall, the analysis methods are appropriate for the data and the analysis seems to have been conducted properly. Though when during analysis some defined re sampling is not used of few observations
 
@@
 
-The results were presented in a well-structured manner. Most tables were well-formated. Maybe Table 1(summary of data) should be made more concise or moved into a supplemental material.
+Overall, the results are adequately presented with tables, the addition of figures and improvements in formatting and labeling could enhance readability and understanding, making the main figures and tables of publication-level quality.
 
@@
 
-* results are very well presented
+* results are presented ok, with room for improvement
 
@@
 
-Discussion is detailed and the interpretation of results is appropriate. Strengths and limitations are specified.
+Overall, the study findings are discussed, there is room for improvement in acknowledging strengths and limitations and interpreting the findings in a more comprehensive and refined manner.
 
@@
 
-* strong, complete and clear discussion
+* minor parts wrong, missing or unclear
 
@@
 
-There are some R scripts showing up in the manuscript.
-And some residual content of the template needs to be removed.
-It would be helpful if the detailed titles/annotations can be added to the figures.
+_The word document still has code when you render it. I think this can be improved and the document will be fine before submission._
 
@@
 
-The project is generally well-structured. All files are in well labeled folders. It is easy to figure out the use for each file/folder.
+The project is well structured, folders are named well, there are few undocumented readme files
 
@@
 
-The R scripts are well-documented with sufficient description.
-Each decision of analysis was justified and explained.
-There are residual content from template in many Rmd files(README).
-The qmd file for supplement material is also from the template.
+The files and folders are labelled well. code is documented well and the output explained but some readme are not well decribed.
 
@@
 
-I was able reproduce the results. Instructions were clear.
+Yes, the results were reproducible, code run smoothly.
 
@@
 
-The research questions have been fully addressed. All sections are relatively complete and thorough.
+Overall, the study is thorough.
 
@@
 
-* strong level of thorougness
+* decent level of thoroughness
 
@@
 
-Overall, this study is in good shape.
-The author may want to consider moving some content into a supplement file.
-Some R scripts show up in the manuscript, so that needs to be fixed.
-The residual content from template in many Rmd files also need to be cleaned.
+_The manuscript needs to be formatted well, put applicable references in the text and improve the discussion but otherwise well done._
